Closed-loop Neuromorphic Benchmarks

### 2015-09-30 paper/paper.tex
changed embodied to closed-loop

\title[Embodied Neuromorphic Benchmarks]{Embodied Neuromorphic Benchmarks}

@@

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is an
    embodied task; that is, a task where the output from the neuromorphic
    hardware affects some environment, which then in turn affects the
    hardware's future input.  However, embodied situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for generating embodied benchmarks that makes use
    of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
    Minimal simulation has been shown to lead to robust real-world performance,
    while still maintaining the practical advantages of simulation, such as
    making it easy for the same benchmark to be used by many researchers.
    This method is flexible enough to allow researchers to explicitly
    modify the benchmarks to identify specific task domains where particular
    hardware excels.  To demonstrate the method, we present a set of novel benchmarks
    that focus on motor control for an arbitrary system with
    unknown external forces.

@@

In this paper, we focus on the development of \emph{embodied} benchmarks.
These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
its own future input} through some environment.  This is in contrast to standard categorization or pattern
identification tasks, where the input is some fixed sequence and the hardware
must produce the correct output for each input (or input pattern).

We believe embodied benchmarks should be of particular interest to
neuromorphic research, given that the most compelling applications of neuromorphic
hardware are likely to be in this domain of embedded and interactive control
of robotic or other physical systems.  However, embodiment itself raises a number
of issues that complicate the development of such benchmarks.  Rather than
simply providing a data file of inputs and desired outputs, the benchmark
must either specify a full physical system for that embodiment, or it must 
provide software for a simulation of that system.  As we discuss below, both
approaches are problematic.  
Describing a method for overcoming these shortcomings is the primary goal of
this paper.

\section{Embodied Benchmarks}

An embodied benchmark task is one where the system we are studying has a 
two-way interaction with some sort of environment.  That is, the outputs from
the neuromorphic hardware are sent to the environment where they cause
an effect, the results of which change the subsequent input.  For example,
the outputs might control the movement of a robot, which in turn affects the
sensory data received by the robot.

@@

To define an embodied benchmark, we need to be explicit about the embodiment.  If
a robot is to be controlled, we need to specify all of the details of that robot.  What motors does it have?  How are they
configured?  How strong are they?  What sensors are there?  Where are they
placed?  How accurate are they?  However, even if these questions are
answered, there is a fundamental problem in that \emph{other researchers
need access to that exact robot}.  If a benchmark is to be widely used, other
researchers developing their own neuromorphic hardware should be able to do their own 
testing on the same benchmark system.

@@

Furthermore, neuromorphic hardware has another constraint that severely limits
the utility of simulation.  Typically, when a simulation is
too simple to reflect reality, we add details to the simulation itself.
Incredibly finely detailed simulations can be created, filling in all of the
details needed.  However, accurate modelling of physical systems can very quickly
become \emph{impractical to run in real time}.
This is a fundamental problem, in that neuromorphic
hardware is often tied to real-time interactions, and there is no way to slow down
the hardware to match the simulated environment.  This means that even if
we spent the considerable amount of research effort needed to define a
simulated environment for an embodied benchmark, running that simulation 
fast enough to interact with the desired hardware would require more
resources than are available.

We are thus left with a situation where any benchmark we might define for
a rich, embodied task will be either impractical for different researchers to
run (if it is physically embodied), or inapplicable to real-world situations
(if it is performed in simulation).  Neither option is particularly appealing.

@@

This approach was first suggested in the context of evolutionary robotics.  
Notably, the problem faced by embodied neuromorphic benchmarking is remarkably 
similar to that faced earlier by these researchers. 
In evolutionary robotics, the
goal is to use genetic algorithms to \emph{evolve} systems that can control
robots to perform various tasks.  These tasks can be as simple as navigation
and obstacle avoidance, but have also included more difficult tasks such as walking,
collecting objects, and visual tracking (\citealt{Nolfi2000}).

@@

To demonstrate this approach to creating embodied neuromorphic benchmarks,
we now consider a basic control task.  Suppose we have a system with a number
of joints with positions $q = [q_1, q_2, ..., q_n]$ and we want to send a control signal $u = [u_1, u_2, ..., u_n]$ to the motors at each joint
such that
the joints move to a particular desired position $q_d = [q_{d,1}, q_{d,2}, ..., q_{d,n}]$.  The only output from the controller is
the signal $u$ and the inputs are the current position
of each motor $q$ and the desired positions $q_d$.

@@

In that case, there are at least two possible ways to implement this algorithm.
First, the multiplication by $d$ can be done on the
output from the neuromorphic hardware.  Any embodied neuromorphic system will have some method that takes
the neural output from the hardware and sends it to the motors (or to the simulation
of the motors).  Instead of sending the result of $Ad$, the hardware could
send $A$ (the activity of all the neurons), and the interface to the motor
can be responsible for doing the multiplication by $d$ and updating $d$
according to the learning rule.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
However, on that benchmark all three systems are implementing exactly 500
neurons.  This demonstrates that the
differences in neuron model across that hardware does not significantly
impact performance.  Given that embodied models rely on real-time simulation,
it is also important to determine how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  With the current implementation, the CPU can
run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
there are 18 on one chip) can run 500
neurons.\footnote{The full chip can run much more than 500 neurons; here we use just a single one of the cores on the SpiNNaker chip.}  These values were measured empirically with the current versions
of the reference Nengo implementation, the Nengo OpenCL implementation,
and the Nengo SpiNNaker implementation (as of August 10, 2015).  All other
parameters are as before.

@@

While it is possible to run large neural models on standard CPUs and GPUs,
one of the primary advantages of neuromorphic hardware is its power
efficiency.  For this reason, the third benchmark normalizes the number of neurons
based on power consumption.  With a power budget of 1W per chip (with 16 used cores),
we estimate 0.0625W for the 500 neurons used here and round up to 0.1W to be
conservative.  Neither the CPU nor the
GPU are designed to run on that little power.  For this reason,
on this benchmark we scale the number of neurons by the power consumption of the hardware.
For this power consumption we measure the difference between the idle power
consumption and the consumption when running the benchmark.  For the Intel
i5-3337U this was $34W - 11.5W = 22.5W$ and for the Nvidia Tesla C2075 GPU
this was $74W - 70W = 4W$.  This value is much lower than the peak power
consumption supported by the GPU ($215W$), indicating that the current implementation
does not make extensive use of the GPU for this task.  Indeed, initial analysis indicates that
the main bottleneck is communication between the GPU and the environment (i.e.,
the minimal simulation), and we feel it is appropriate that this benchmark
captures that limitation of the current GPU implementation.  The GPU could easily
run many more neurons than this in real time, if those neurons were not connected
to an environment.  However, that would not be useful for an embodied task.

>>>>

\title[Closed-loop Neuromorphic Benchmarks]{Closed-loop Neuromorphic Benchmarks}

@@

    Evaluating the effectiveness and performance of neuromorphic hardware is
    difficult.  It is even more difficult when the task of interest is a
    closed-loop task; that is, a task where the output from the neuromorphic
    hardware affects some environment, which then in turn affects the
    hardware's future input.  However, closed-loop situations
    are one of the primary potential uses of neuromorphic hardware.  To address
    this, we present a methodology for generating closed-loop benchmarks that makes use
    of a hybrid of real physical embodiment and a type of ``minimal'' simulation.
    Minimal simulation has been shown to lead to robust real-world performance,
    while still maintaining the practical advantages of simulation, such as
    making it easy for the same benchmark to be used by many researchers.
    This method is flexible enough to allow researchers to explicitly
    modify the benchmarks to identify specific task domains where particular
    hardware excels.  To demonstrate the method, we present a set of novel benchmarks
    that focus on motor control for an arbitrary system with
    unknown external forces.

@@

In this paper, we focus on the development of \emph{closed-loop} benchmarks.
These are dynamic tasks where the output of the neuromorphic hardware \emph{influences
its own future input} through some environment.  This is in contrast to standard categorization or pattern
identification tasks, where the input is some fixed sequence and the hardware
must produce the correct output for each input (or input pattern).

We believe closed-loop benchmarks should be of particular interest to
neuromorphic research, given that the most compelling applications of neuromorphic
hardware are likely to be in this domain of embedded and interactive control
of robotic or other physical systems.  However, the closed loop itself raises a number
of issues that complicate the development of such benchmarks.  Rather than
simply providing a data file of inputs and desired outputs, the benchmark
must either specify a full physical system to be controlled, or it must 
provide software for a simulation of that system.  As we discuss below, both
approaches are problematic.  
Describing a method for overcoming these shortcomings is the primary goal of
this paper.

\section{Closed-Loop Benchmarks}

A closed-loop benchmark task is one where the system we are studying has a 
two-way interaction with some sort of environment.  That is, the outputs from
the neuromorphic hardware are sent to the environment where they cause
an effect, the results of which change the subsequent input.  For example,
the outputs might control the movement of a robot, which in turn affects the
sensory data received by the robot.

@@

To define a closed-loop benchmark, we need to be explicit about the interaction with the environment.  If
a robot is to be controlled, we need to specify all of the details of that robot.  What motors does it have?  How are they
configured?  How strong are they?  What sensors are there?  Where are they
placed?  How accurate are they?  However, even if these questions are
answered, there is a fundamental problem in that \emph{other researchers
need access to that exact robot}.  If a benchmark is to be widely used, other
researchers developing their own neuromorphic hardware should be able to do their own 
testing on the same benchmark system.

@@

Furthermore, neuromorphic hardware has another constraint that severely limits
the utility of simulation.  Typically, when a simulation is
too simple to reflect reality, we add details to the simulation itself.
Incredibly finely detailed simulations can be created, filling in all of the
details needed.  However, accurate modelling of physical systems can very quickly
become \emph{impractical to run in real time}.
This is a fundamental problem, in that neuromorphic
hardware is often tied to real-time interactions, and there is no way to slow down
the hardware to match the simulated environment.  This means that even if
we spent the considerable amount of research effort needed to define a
simulated environment for a closed-loop benchmark, running that simulation 
fast enough to interact with the desired hardware would require more
resources than are available.

We are thus left with a situation where any benchmark we might define for
a closed-loop task will be either impractical for different researchers to
run (if it is physically embodied), or inapplicable to real-world situations
(if it is performed in simulation).  Neither option is particularly appealing.

@@

This approach was first suggested in the context of evolutionary robotics.  
Notably, the problem faced by closed-loop neuromorphic benchmarking is remarkably 
similar to that faced earlier by these researchers. 
In evolutionary robotics, the
goal is to use genetic algorithms to \emph{evolve} systems that can control
robots to perform various tasks.  These tasks can be as simple as navigation
and obstacle avoidance, but have also included more difficult tasks such as walking,
collecting objects, and visual tracking (\citealt{Nolfi2000}).

@@

To demonstrate this approach to creating closed-loop neuromorphic benchmarks,
we now consider a basic control task.  Suppose we have a system with a number
of joints with positions $q = [q_1, q_2, ..., q_n]$ and we want to send a control signal $u = [u_1, u_2, ..., u_n]$ to the motors at each joint
such that
the joints move to a particular desired position $q_d = [q_{d,1}, q_{d,2}, ..., q_{d,n}]$.  The only output from the controller is
the signal $u$ and the inputs are the current position
of each motor $q$ and the desired positions $q_d$.

@@

In that case, there are at least two possible ways to implement this algorithm.
First, the multiplication by $d$ can be done on the
output from the neuromorphic hardware.  Any closed-loop neuromorphic system will have some method that takes
the neural output from the hardware and sends it to the motors (or to the simulation
of the motors).  Instead of sending the result of $Ad$, the hardware could
send $A$ (the activity of all the neurons), and the interface to the motor
can be responsible for doing the multiplication by $d$ and updating $d$
according to the learning rule.

@@

In Figure \ref{fig:analysis_basic}, all three systems perform equally well.
However, on that benchmark all three systems are implementing exactly 500
neurons.  This demonstrates that the
differences in neuron model across that hardware does not significantly
impact performance.  Given that closed-loop models rely on real-time simulation,
it is also important to determine how many neurons each piece of hardware is capable of running in real time.
This is shown in Figure \ref{fig:analysis_compute}.  With the current implementation, the CPU can
run 5200 neurons, the GPU 1500 neurons, and a single SpiNNaker core (of which
there are 18 on one chip) can run 500
neurons.\footnote{The full chip can run much more than 500 neurons; here we use just a single one of the cores on the SpiNNaker chip.}  These values were measured empirically with the current versions
of the reference Nengo implementation, the Nengo OpenCL implementation,
and the Nengo SpiNNaker implementation (as of August 10, 2015).  All other
parameters are as before.

@@

While it is possible to run large neural models on standard CPUs and GPUs,
one of the primary advantages of neuromorphic hardware is its power
efficiency.  For this reason, the third benchmark normalizes the number of neurons
based on power consumption.  With a power budget of 1W per chip (with 16 used cores),
we estimate 0.0625W for the 500 neurons used here and round up to 0.1W to be
conservative.  Neither the CPU nor the
GPU are designed to run on that little power.  For this reason,
on this benchmark we scale the number of neurons by the power consumption of the hardware.
For this power consumption we measure the difference between the idle power
consumption and the consumption when running the benchmark.  For the Intel
i5-3337U this was $34W - 11.5W = 22.5W$ and for the Nvidia Tesla C2075 GPU
this was $74W - 70W = 4W$.  This value is much lower than the peak power
consumption supported by the GPU ($215W$), indicating that the current implementation
does not make extensive use of the GPU for this task.  Indeed, initial analysis indicates that
the main bottleneck is communication between the GPU and the environment (i.e.,
the minimal simulation), and we feel it is appropriate that this benchmark
captures that limitation of the current GPU implementation.  The GPU could easily
run many more neurons than this in real time, if those neurons were not connected
to an environment.  However, that would not be useful for a closed-loop task.
